An Empirical Analysis of the Role of Energy in Economic Growth

### 2015-02-01 Articles/Article_EnerEcon_2014/EconGrowth4.Rnw
Added identifying number to the boundary mod tbl.
Changed order of rows in the boundary mod table.

In addition to the above trials, 
and in a manner similar to our approach to the Cobb-Douglas model,
we fit along all possible CES boundaries.
Boundaries are given by all combinations of the following:
$\delta_1$~=~0~or~1, 
$\sigma_1$~=~0~or~$\infty$,
$\delta$~=~0~or~1, and
$\sigma$~=~0~or~$\infty$.
Equation~\ref{eq:ces_generic} shows the generic three-factor CES equation,
wherein factors of production ($x_1$,~$x_2$,~$x_3$) 
are permutations of ($k$,~$l$,~$e$).
%
\begin{equation} \label{eq:ces_generic}
  y = \gamma A \left\{ \delta \left[ \delta_1 x_1^{-\rho_1} 
      + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{\rho/\rho_1}  
          + \left( 1 - \delta \right) x_3^{-\rho}  \right\} ^ {-1/\rho}
\end{equation}
%
Table~\ref{tab:CES_boundary_models} shows the set 
of 20 degenerate equations found along parameter boundaries.
Each degenerate equation provides a boundary model.
Note that many boundary parameter combinations 
yield equivalent CES boundary models,%
  \footnote{
  For example, all possible boundary parameter combinations 
  that include $\delta = 0$ yield the same boundary model, $y = \gamma A x_3$.
  }
so there are far fewer than the maximum possible number of degenerate equations~(81).%
  \footnote{
  There are three possible states (unspecified, lower boundary, upper boundary)
  for four parameters ($\delta_1$, $\sigma_1$, $\delta$, $\sigma$),
  which gives $3^4 = 81$ possible boundary models.
  }
Boundary models obtained when $\delta_1$ or $\delta$ is at an extreme value
are straightforward to derive from Equation~\ref{eq:ces_generic}.
When $\sigma_1$ or $\sigma$ is 0, 
factors of production are complimentary,
and the Leontief model applies.%
  \footnote{
  For the case of $\sigma_1 = 0$ and $\delta~=~1$, the Leontief model is
  $y = \gamma A \min (x_1, x_2)$.
  }
When $\sigma_1$ or $\sigma$ is $\infty$, 
factors of production are subsitutible,
and the linear model applies.%
  \footnote{
  For the case of $\sigma_1 = \infty$ and $\delta~=~1$, the linear model is
  $y = \gamma A \left[ \delta_1 x_1 + (1 - \delta_1) x_2 \right]$.
  }
In the parameter columns of Table~\ref{tab:CES_boundary_models}, 
``---'' indicates that a parameter is unknowable on that boundary
(because it does not appear in the boundary model), and
an empty space indicates that the parameter 
can be estimated on that boundary using the degenerate equation.
The top portion of Table~\ref{tab:CES_boundary_models} shows 
two-parameter CES boundary models
(similar to the CES model without energy in Equation~\ref{eq:CES})
with $\delta~=~1$, $\sigma$~unknowable, and $x_3$~absent.
If the fitted parameters in a boundary model violate constraints, 
the boundary model is rejected.

\begin{table} \caption{Boundary models for the CES equation.} 
\label{tab:CES_boundary_models} 
  \begin{center}
    \begin{tabular}{c c c c l} 
      \toprule
      $\delta_1$  & $\sigma_1$ & $\delta$   & $\sigma$ & Boundary model \\
      \midrule
      1           & ---        & 1          & ---      & $y = \gamma A x_1$ \\
      0           & ---        & 1          & ---      & $y = \gamma A x_2$ \\
      ---         & 0          & 1          & ---      & $y = \gamma A \min (x_1, \, x_2)$ \\
%
                  & $\infty$   & 1          & ---        
            & $y = \gamma A \left[ \delta_1 x_1 + (1 - \delta_1) x_2 \right]$ \\
%
                  &            & 1          & ---
            & $y = \gamma A \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{-1/\rho_1}$ \\
\midrule
      ---         & ---        & 0          & ---      & $y = \gamma A x_3$ \\
      1           & ---        & ---        & 0        & $y = \gamma A \min (x_1, \, x_3)$ \\
      0           & ---        & ---        & 0        & $y = \gamma A \min (x_2, \, x_3)$ \\
      ---         & 0          & ---        & 0        & $y = \gamma A \min (x_1, \, x_2, \, x_3)$ \\
%
      1           & ---        &            & $\infty$        
            & $y = \gamma A \left[ \delta x_1 + (1 - \delta) x_3 \right]$ \\
%
      0           & ---        &            & $\infty$        
            & $y = \gamma A \left[ \delta x_2 + (1 - \delta) x_3 \right]$ \\
%
                  & $\infty$   &            & $\infty$        
            & $y = \gamma A \left\{ \delta \left[ \delta_1 x_1 + \left( 1 - \delta_1 \right) x_2 \right] 
                    + \left( 1 - \delta \right) x_3  \right\}$ \\
%
      ---         & 0          &            & $\infty$        
            & $y = \gamma A \left[ \delta \min (x_1, \, x_2) + \left( 1 - \delta \right) x_3  \right]$ \\
%
                  & $\infty$   & ---        & 0
            & $y = \gamma A \min \left[ \delta_1 x_1 + \left(1 - \delta_1 \right) x_2, \, x_3  \right]$ \\
%
      1           & ---        &         & 
            & $y = \gamma A \left[ \delta x_1^{-\rho} 
                    + \left(1 - \delta \right) x_3^{-\rho} \right] ^ {-1/\rho}$ \\ 
%
      0           & ---        &            & 
            & $y = \gamma A \left[ \delta x_2^{-\rho} + \left(1 - \delta \right) x_3^{-\rho} \right]^{-1/\rho}$ \\
%
                  &            & ---     & 0
            & $y = \gamma A \min \left\{ \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{-1/\rho_1}, \, x_3 \right\}$ \\ 
%
      ---         & 0          &            & 
            & $y = \gamma A \left\{ \delta \left[ \min \left( x_1, \, x_2 \right) \right]^{-\rho}  
                    + \left( 1 - \delta \right) x_3^{-\rho}  \right\} ^ {-1/\rho}$ \\ 
%
                  &            &            & $\infty$
            & $y = \gamma A \left\{ \delta \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{-1/\rho_1}  
                    + \left( 1 - \delta \right) x_3  \right\}$ \\ 
%
                  & $\infty$   &            &
            & $y = \gamma A \left\{ \delta \left[ \delta_1 x_1 
                    + \left( 1 - \delta_1 \right) x_2 \right]^{-\rho}  
                    + \left( 1 - \delta \right) x_3^{-\rho}  \right\} ^ {-1/\rho}$ \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

>>>>

In addition to the above trials, 
and in a manner similar to our approach to the Cobb-Douglas model,
we fit along all possible CES boundaries.
Boundaries are given by all combinations of the following:
$\delta_1$~=~0~or~1, 
$\sigma_1$~=~0~or~$\infty$,
$\delta$~=~0~or~1, and
$\sigma$~=~0~or~$\infty$.
Equation~\ref{eq:ces_generic} shows the generic three-factor CES equation,
wherein factors of production ($x_1$,~$x_2$,~$x_3$) 
are permutations of ($k$,~$l$,~$e$).
%
\begin{equation} \label{eq:ces_generic}
  y = \gamma A \left\{ \delta \left[ \delta_1 x_1^{-\rho_1} 
      + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{\rho/\rho_1}  
          + \left( 1 - \delta \right) x_3^{-\rho}  \right\} ^ {-1/\rho}
\end{equation}
%
Table~\ref{tab:CES_boundary_models} shows the set 
of 20 degenerate equations found along parameter boundaries.
Each degenerate equation provides a boundary model.
Note that many boundary parameter combinations 
yield equivalent CES boundary models,%
  \footnote{
  For example, all possible boundary parameter combinations 
  that include $\delta = 0$ yield the same boundary model, $y = \gamma A x_3$.
  }
so there are far fewer than the maximum possible number of degenerate equations~(81).%
  \footnote{
  There are three possible states (unspecified, lower boundary, upper boundary)
  for four parameters ($\delta_1$, $\sigma_1$, $\delta$, $\sigma$),
  which gives $3^4 = 81$ possible boundary models.
  }
Boundary models obtained when $\delta_1$ or $\delta$ is at an extreme value
are straightforward to derive from Equation~\ref{eq:ces_generic}.
When $\sigma_1$ or $\sigma$ is 0, 
factors of production are complimentary,
and the Leontief model applies.%
  \footnote{
  E.g., when $\sigma_1 = 0$ and $\delta~=~1$, the Leontief model is
  $y = \gamma A \min (x_1, x_2)$.
  }
When $\sigma_1$ or $\sigma$ is $\infty$, 
factors of production are subsitutible,
and the linear model applies.%
  \footnote{
  E.g., when $\sigma_1 = \infty$ and $\delta~=~1$, the linear model is
  $y = \gamma A \left[ \delta_1 x_1 + (1 - \delta_1) x_2 \right]$.
  }
In the parameter columns of Table~\ref{tab:CES_boundary_models}, 
``---'' indicates that a parameter is unknowable on that boundary
(because it does not appear in the boundary model), and
an empty space indicates that the parameter 
can be estimated on that boundary using the degenerate equation.
The top portion of Table~\ref{tab:CES_boundary_models} shows 
two-parameter CES boundary models
(similar to the CES model without energy in Equation~\ref{eq:CES})
with $\delta~=~1$, $\sigma$~unknowable, and $x_3$~absent.
If the fitted parameters in a boundary model violate constraints, 
the boundary model is rejected.

\begin{table} \caption{Boundary models for the CES equation.} 
\label{tab:CES_boundary_models} 
  \begin{center}
    \begin{tabular}{r c c c c l} 
      \toprule
      & $\delta_1$  & $\sigma_1$ & $\delta$   & $\sigma$ & Boundary model \\
      \midrule
      1. & 1           & ---        & 1          & ---      & $y = \gamma A x_1$ \\
      2. & 0           & ---        & 1          & ---      & $y = \gamma A x_2$ \\
      3. & ---         & 0          & 1          & ---      & $y = \gamma A \min (x_1, \, x_2)$ \\
%
      4. &             & $\infty$   & 1          & ---        
            & $y = \gamma A \left[ \delta_1 x_1 + (1 - \delta_1) x_2 \right]$ \\
%
      5. &             &            & 1          & ---
            & $y = \gamma A \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{-1/\rho_1}$ \\
\midrule
      6. & ---         & ---        & 0          & ---      & $y = \gamma A x_3$ \\
      7. & 1           & ---        & ---        & 0        & $y = \gamma A \min (x_1, \, x_3)$ \\
      8. & 0           & ---        & ---        & 0        & $y = \gamma A \min (x_2, \, x_3)$ \\
      9. & ---         & 0          & ---        & 0        & $y = \gamma A \min (x_1, \, x_2, \, x_3)$ \\
%
     10. & 1           & ---        &            & $\infty$        
            & $y = \gamma A \left[ \delta x_1 + (1 - \delta) x_3 \right]$ \\
%
     11. & 0           & ---        &            & $\infty$        
            & $y = \gamma A \left[ \delta x_2 + (1 - \delta) x_3 \right]$ \\
%
     12. &             & $\infty$   &            & $\infty$        
            & $y = \gamma A \left\{ \delta \left[ \delta_1 x_1 + \left( 1 - \delta_1 \right) x_2 \right] 
                    + \left( 1 - \delta \right) x_3  \right\}$ \\
%
     13. & ---         & 0          &            & $\infty$        
            & $y = \gamma A \left[ \delta \min (x_1, \, x_2) + \left( 1 - \delta \right) x_3  \right]$ \\
%
     14. &             & $\infty$   & ---        & 0
            & $y = \gamma A \min \left[ \delta_1 x_1 + \left(1 - \delta_1 \right) x_2, \, x_3  \right]$ \\
%
     15. & 1           & ---        &         & 
            & $y = \gamma A \left[ \delta x_1^{-\rho} 
                    + \left(1 - \delta \right) x_3^{-\rho} \right] ^ {-1/\rho}$ \\ 
%
     16. & 0           & ---        &            & 
            & $y = \gamma A \left[ \delta x_2^{-\rho} + \left(1 - \delta \right) x_3^{-\rho} \right]^{-1/\rho}$ \\
%
     17. &             & $\infty$   &            &
            & $y = \gamma A \left\{ \delta \left[ \delta_1 x_1 
                    + \left( 1 - \delta_1 \right) x_2 \right]^{-\rho}  
                    + \left( 1 - \delta \right) x_3^{-\rho}  \right\} ^ {-1/\rho}$ \\
%
     18. & ---         & 0          &            & 
            & $y = \gamma A \left\{ \delta \left[ \min \left( x_1, \, x_2 \right) \right]^{-\rho}  
                    + \left( 1 - \delta \right) x_3^{-\rho}  \right\} ^ {-1/\rho}$ \\ 
%
     19. &             &            &            & $\infty$
            & $y = \gamma A \left\{ \delta \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{-1/\rho_1}  
                    + \left( 1 - \delta \right) x_3  \right\}$ \\ 
%
     20. &            &            & ---     & 0
            & $y = \gamma A \min \left\{ \left[ \delta_1 x_1^{-\rho_1} 
                    + \left( 1 - \delta_1 \right) x_2^{-\rho_1} \right]^{-1/\rho_1}, \, x_3 \right\}$ \\ 
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}
